---
title: "Lab 9 pt 1"
author: "Max Settineri"
date: "2023-03-15"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(broom)

# Time series packages
library(tsibble)
library(feasts)
library(fable)

```

## Read in energy data and convert to a tsibble

Data from US EIA on US residential energy consumption from Jan 1973 - Oct 2017. Units in trillions of BTUs

```{r}
energy_df <- read_csv(here("data/energy.csv"))

energy_ts <- energy_df %>% 
  mutate(month_new = lubridate::ym(month)) %>% ## workaround to avoid weird tsibble error 
  mutate(date = tsibble::yearmonth(month_new)) %>% 
  as_tsibble(key = NULL, index = date)
## head(energy_ts) # can check data type for each column
```

## Exploratort time series visualization

```{r}
ggplot(data = energy_ts, aes(x = date, y = res_total)) +
  geom_line() +
  labs(y = "Residential energy consumption \n (Trillion BTU)")
```

### season plot

```{r}
energy_ts %>% 
  gg_season(y = res_total) +
  theme_minimal() +
  scale_color_viridis_c() +
  labs(x = "month",
       y = "residential energy consumption (trillion BTU)")
```

## subseries plot

```{r}
energy_ts %>% 
  gg_subseries(y = res_total)

```

## Decomposition

STL decomposition vs. classical decomposition. Classical has same seasonality for entire time series, but STL allows seasonality to change over time.
```{r}
## calculate STL decomposition
dcmp <- energy_ts %>% 
  model(STL(res_total ~ season()))
### components(dcmp)

## visualize
components(dcmp) %>% 
  autoplot() +
  theme_minimal()
```

## Autocorrelation function (ACF)

```{r}
energy_ts %>% 
  ACF(y = res_total) %>% 
  autoplot() 
```

## Forecasting - Holt-winters exponential smoothing

```{r}
## create our model
energy_fit <- energy_ts %>% 
  model(ets = ETS(res_total ~ season('M'))) 

## create forecast based on our model 10 years into the future
energy_forecast <- energy_fit %>% 
  forecast(h = "10 years")

energy_forecast %>%
  autoplot()
```

Now plot actual res_total values on top of the predicted values

```{r}
## append predicted values (and residuals) to the original energy data
energy_predicted <- broom::augment(energy_fit)

ggplot(data = energy_predicted) +
  geom_line(aes(x = date, y = res_total), color = 'black') +
  geom_line(aes(x = date, y = .fitted), color = 'red', alpha = 0.7)

ggplot(data = energy_predicted, aes(x = .resid)) +
  geom_histogram()

ggplot(data = energy_predicted, aes(x = date, y = .resid)) +
  geom_point()
```

