---
title: "lab 9 part 3"
author: "Max Settineri"
date: "2023-03-15"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(tidytext)
library(textdata)
library(pdftools)
library(ggwordcloud)
```

```{r}
hobbit_text <- pdf_text(here::here('data', 'the-hobbit.pdf'))

hobbit_p34 <- hobbit_text[34]
```


## Get text into a dataframe

```{r}
hobbit_lines <- data.frame(hobbit_text) %>% 
  mutate(page = 1:n()) %>%
  mutate(text_full = str_split(hobbit_text, pattern = '\n')) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full))
```

## Doing some tidying

```{r}
hobbit_chapts <- hobbit_lines %>% 
  slice(-(1:137)) %>% 
  mutate(chapter = ifelse(str_detect(text_full, "Chapter"), text_full, NA)) %>% 
  fill(chapter, .direction = 'down') %>% 
  separate(col = chapter, into = c("ch", "num"), sep = " ") %>% 
  mutate(chapter = as.numeric(as.roman(num)))
```

Get some word counts by chapter

```{r}
hobbit_words <- hobbit_chapts %>% 
  unnest_tokens(word, text_full, token = 'words') 

hobbit_wordcount <- hobbit_words %>% 
  # group_by(chapter, word) %>% 
  # summarize(n = n())
  count(chapter, word)
```

## Remove stop words

```{r}
x <- stop_words

hobbit_words_clean <- hobbit_words %>% 
  anti_join(stop_words, by = 'word')

non_stop_counts <- hobbit_words_clean %>% 
  count(chapter, word)
```

## Find the top 5 words from each chapter

```{r}
top_5_words <- non_stop_counts %>% 
  group_by(chapter) %>% 
  slice_max(order_by = n, n = 5)

ggplot(data = top_5_words, aes(x = n, y = word)) +
  geom_col(fill = "blue") +
  facet_wrap(~ chapter, scales = "free")
```

## Make a word cloud for chapter 1

```{r}
ch1_top100 <- non_stop_counts %>% 
  filter(chapter == 1) %>% 
  slice_max(order_by = n, n = 100)

ch1_cloud <- ggplot(data = ch1_top100, aes(label = word)) +
  geom_text_wordcloud(aes(color = n, size = n), shape = 'diamond') +
  scale_size_area(max_size = 6) +
  scale_color_gradientn(colors = c('darkgreen', 'blue', 'purple'))

ch1_cloud
```

## Sentiment analysis

### afinn lexicon

Ranks each word in lexicon on scale from -5 to +5

```{r}
afinn_lex <- get_sentiments(lexicon = "afinn")

afinn_pos <- get_sentiments(lexicon = 'afinn') %>% 
  filter(value %in% c(3, 4, 5))

DT::datatable(afinn_pos)

```

### bing lexiconn

Binary scoring of words, either positive or negative

```{r}
bing_lex <- get_sentiments(lexicon = 'bing')
```

### nrc lexicon

NRC lexicon categorizes each word into bins

```{r}
nrc_lexicon <- get_sentiments(lexicon = 'nrc')
```

### Sentiment analysis with afinn

First, inner_join the lexicon to the word dataframe

```{r}
hobbit_afinn <- hobbit_words_clean %>% 
  inner_join(afinn_lex, by = 'word')

afinn_counts <- hobbit_afinn %>% 
  group_by(chapter, value) %>% 
  summarize(n = n())

ggplot(afinn_counts, aes(x = value, y = n)) +
  geom_col() +
  facet_wrap(~chapter)

afinn_means <- hobbit_afinn %>% 
  group_by(chapter) %>% 
  summarize(mean_afinn = mean(value))

ggplot(data = afinn_means,
       aes(x = mean_afinn, y = fct_rev(factor(chapter)))) +
  geom_col() +
  labs(y = 'chapter')
```

### Sentiment analysis with bing

```{r}
hobbit_bing <- hobbit_words_clean %>% 
  inner_join(bing_lex, by = 'word')

bing_counts <- hobbit_bing %>% 
  count(chapter, sentiment)

ggplot(bing_counts, aes(x = sentiment, y = n)) +
  geom_col() +
  facet_wrap(~ chapter)

### find log positive-to-negative ratio
bing_log_ratio_book <- hobbit_bing %>% 
  summarize(n_pos = sum(sentiment == 'positive'),
            n_neg = sum(sentiment == 'negative'),
            log_ratio = log(n_pos / n_neg))

bing_log_ratio_ch <- hobbit_bing %>% 
  group_by(chapter) %>% 
  summarize(n_pos = sum(sentiment == 'positive'),
            n_neg = sum(sentiment == 'negative'),
            log_ratio = log(n_pos / n_neg)) %>% 
  mutate(log_ratio_adjust = log_ratio - bing_log_ratio_book$log_ratio) %>% 
  mutate(pos_neg = ifelse(log_ratio_adjust > 0, 'pos', 'neg'))

ggplot(data = bing_log_ratio_ch,
       aes(x = log_ratio_adjust,
           y = fct_rev(factor(chapter)),
           fill = pos_neg)) +
  geom_col() +
  labs(x = "adjusted log(pos/neg)",
       y = "chapter") +
  scale_fill_manual(values = c('pos' = 'slateblue', 'neg' = 'darkred')) +
  theme_minimal() +
  theme(legend.position = 'none')
```

### Sentiment analysis with NRC lexicon

```{r}
hobbit_nrc <- hobbit_words_clean %>% 
  inner_join(nrc_lexicon, by = 'word')

hobbit_nrc_counts <- hobbit_nrc %>% 
  count(chapter, sentiment)

ggplot(data = hobbit_nrc_counts, aes(x = n, y = sentiment)) +
  geom_col() +
  facet_wrap(~ chapter)

ggplot(data = hobbit_nrc_counts, aes(x = n, y = factor(chapter) %>% fct_rev())) +
  geom_col() +
  facet_wrap(~ sentiment) +
  labs(y = 'chapter')
```

